Evaluating Staffing Characteristics, Healthcare-associated Infections, and Provider Characteristics in U.S. Nursing Homes, 2021

### 2023-03-18 code/analysis_code/exploratory_analysis.qmd
Part 3

```{r}
#load needed packages. make sure they are installed.
library(here) #for data loading/saving
library(dplyr)
library(skimr)
library(ggplot2)
```

@@

First, Risk Standardized HAI Rate
```{r}
p1 <- mydata %>% ggplot(aes(x=HAI_039_RS_Rate)) + geom_histogram() 
plot(p1)
figure_file = here("results","RiskStandardized_HAI_Rate_distribution.png")
ggsave(filename = figure_file, plot=p1)

@@

```{r}
p3 <- mydata %>% ggplot(aes(x=Adjusted.Total.Nurse.Staffing.Hours.per.Resident.per.Day, y=HAI_039_RS_Rate, color=Staffing.Rating)) + geom_point() + geom_smooth(method='lm')
plot(p3)
figure_file = here("results","staff_hai.png")
ggsave(filename = figure_file, plot=p3) 
```

@@

```{r}
p4 <- mydata %>% ggplot(aes(x=Ownership.Type, y=HAI_039_RS_Rate)) + geom_boxplot() + geom_smooth(method='lm')
plot(p4)
figure_file = here("results","staff_hai_ownership.png")
ggsave(filename = figure_file, plot=p4) 
```

# Notes

Might want to regroup ownership type into fewer categories (e.g., put all government and for-profit together.

>>>>

```{r}
#load needed packages. make sure they are installed.
library(here) 
library (tidyverse)

library(skimr)

library(ggplot2)
```

@@

First, Risk Standardized HAI Rate

```{r}
p1 <- mydata %>% ggplot(aes(x=HAI_039_RS_Rate)) + geom_histogram() 
plot(p1)
figure_file = here("results","RiskStandardized_HAI_Rate_distribution.png")
ggsave(filename = figure_file, plot=p1)

@@

```{r}
p3 <- mydata %>% ggplot(aes(x=Adjusted.Total.Nurse.Staffing.Hours.per.Resident.per.Day, y=HAI_039_RS_Rate, color=Staffing.Rating)) + geom_point() + geom_smooth(method='lm')
plot(p3)
figure_file = here("results","staff_hai.png")
ggsave(filename = figure_file, plot=p3)

p3b <- mydata %>% ggplot(aes(x=as.factor(Staffing.Rating), y=HAI_039_RS_Rate)) + geom_boxplot() + geom_jitter(width=0.15, alpha=0.08, color="black", fill="white")
plot(p3b)

figure_file = here("results","staffrating_hai.png")
ggsave(filename = figure_file, plot=p3b)

@@

```{r}
p4 <- mydata %>% ggplot(aes(x=as.factor(Ownership.Type), y=HAI_039_RS_Rate)) + geom_boxplot() + geom_smooth(method='lm')
plot(p4)
figure_file = here("results","staff_hai_ownership.png")
ggsave(filename = figure_file, plot=p4) 
```

# Changing our outcome

In this plot we categorized our outcome as a 3-level variable: better than the national rate, worse than the national rate or no different than the national rate.

```{r}

p5 <-  mydata %>% ggplot(aes(x=as.factor(COMP_PERF), y=Adjusted.Total.Nurse.Staffing.Hours.per.Resident.per.Day)) + geom_boxplot() + geom_jitter(width=0.15, alpha=0.08, color="black", fill="white")
plot(p5)

figure_file = here("results","staff_hai_compscore.png")
ggsave(filename = figure_file, plot=p5)

We find that while nursing homes who have rates better than the national rate do appear to have higher staffing scores, nursing homes no different and worse than the national rate appear more similar in distribution.

## Modeling Fun

For this section I have decided I am going to assess if staffing rating adds additional information to my model that includes staffing rate. For this model we will use the risk adjusted HAI rate as a continuous outcome. We will use a linear model. We will make three models:

1.  Model with staffing rate and other covariates of interest (Ownership Type, Average Number of Residents per Day, and staffing rate (the continuous value)
2.  Model with staffing rating and other covariates of interest (Ownership Type, Average Number of Residents per day, and staffing rating (the categorical value)
3.  Model with both staffing rating and staffing rate + other covariates of interest

```{r}
#Model 1

model1 <- lm(HAI_039_RS_Rate ~ as.factor(Ownership.Type) + Average.Number.of.Residents.per.Day + Adjusted.Total.Nurse.Staffing.Hours.per.Resident.per.Day + Number.of.Citations.from.Infection.Control.Inspections + Total.Weighted.Health.Survey.Score, data=mydata)

summary(model1)

model2 <- lm(HAI_039_RS_Rate ~ as.factor(Ownership.Type) + Average.Number.of.Residents.per.Day + as.factor(Staffing.Rating)+ Number.of.Citations.from.Infection.Control.Inspections + Total.Weighted.Health.Survey.Score, data=mydata)

summary(model2)

model3 <- lm(HAI_039_RS_Rate ~ as.factor(Ownership.Type) + Average.Number.of.Residents.per.Day + Adjusted.Total.Nurse.Staffing.Hours.per.Resident.per.Day+ as.factor(Staffing.Rating) + Number.of.Citations.from.Infection.Control.Inspections + Total.Weighted.Health.Survey.Score, data=mydata)

summary(model3)

My key takeaways from these three models is that Model 2 and Model 3 have very similar R\^2 fits. So they are likely not adding that much information by including both staffing rating and the staffing rate in the same model. Also, while less important, we see that staffing rate is not significantly associated with our outcome in this fully adjusted model.

I think currently I am leaning towards choosing model 2 as a our final model for associations. This will allow us to draw conclusions about how staffing ratings are associated with healthcare associated infection rates.

### 2023-03-18 code/processing_code/processingfile_v1.qmd
Part 3

```{r}
raw_provider1 <- raw_provider[c("Federal.Provider.Number", "Average.Number.of.Residents.per.Day",
                   "Adjusted.Total.Nurse.Staffing.Hours.per.Resident.per.Day",
                   "Staffing.Rating", "Provider.State",   "Ownership.Type")]
skim(raw_provider1)

@@

raw_quality1 <- raw_quality %>% 
  select (cms_certification_number_ccn, state, HAI_039_Observed, HAI_039_Obs_Rate, HAI_039_Denominator, HAI_039_RS_Rate)

>>>>

```{r}
raw_provider1 <- raw_provider[c("Federal.Provider.Number", "Average.Number.of.Residents.per.Day",
                   "Adjusted.Total.Nurse.Staffing.Hours.per.Resident.per.Day",
                   "Staffing.Rating", "Provider.State",   "Ownership.Type", 
                   "Total.Weighted.Health.Survey.Score", "Number.of.Citations.from.Infection.Control.Inspections")]
skim(raw_provider1)

@@

raw_quality1 <- raw_quality %>% 
  select (cms_certification_number_ccn, state, HAI_039_Observed, HAI_039_Obs_Rate, HAI_039_Denominator, HAI_039_RS_Rate, COMP_PERF)

### 2023-03-18 products/manuscript/Manuscript.qmd
Part 3

Part 1: Project Overview

@@

      The prposed study aims to assess various nursing home characteristics for association with healthcare-associated infection rates.

      First, we will provide a descriptive analysis of nursing homes across the United States including location, bed size, ownership, average number of residents per day. Specifically of interest, we will also describe variability in overall quality ratings, the healthcare-associated infection quality reporting measure, and measures of staffing (including various levels of nursing care (i.e., RN, LPN, and aide)  both aggregated and separately). 
      Second, we will utilize ANOVA methods to determine which, if any,  average staffing characteristics are associated with the classifications of HAI rates as “better”, “worse”, or “as expected”.  
      Third, we will assess a variety of nursing home characteristics and selected staffing measures for associations with  the risk adjusted rate of HAIs in the nursing home.

# Summary/Abstract
_Write a summary of your project._

@@

## General Background Information
_Provide enough background on your topic that others can understand the why and how of your analysis_

## Description of data and data source
_Describe what the data is, what it contains, where it is from, etc. Eventually this might be part of a methods section._

## Questions/Hypotheses to be addressed
_State the research questions you plan to answer with this analysis._

@@

_Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement._

## Data aquisition
_As applicable, explain where and how you got the data. If you directly import the data from an online source, you can combine this section with the next._

## Data import and cleaning
_Write code that reads in the file and cleans it so it's ready for analysis. Since this will be fairly long code for most datasets, it might be a good idea to have it in one or several R scripts. If that is the case, explain here briefly what kind of cleaning/processing you do, and provide more details and well documented code somewhere (e.g. as supplement in a paper). All materials, including files that contain code, should be commented well so everyone can follow along._

## Statistical analysis
_Explain anything related to your statistical analyses._

@@

_Use a combination of text/tables/figures to explore and describe your data. Show the most important descriptive results here. Additional ones should go in the supplement. Even more can be in the R and Quarto files that are part of your project._

@@

_To get some further insight into your data, if reasonable you could compute simple statistics (e.g. simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable. Though note that unless you pre-specified the outcome and main exposure, any "p<0.05 means statistical significance" interpretation is not valid._

@@

_Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc. This might again be code that is best placed in one or several separate R scripts that need to be well documented. You want the code to produce figures and data ready for display as tables, and save those. Then you load them here._

@@

## Summary and Interpretation
_Summarize what you did, what you found and what it means._

## Strengths and Limitations
_Discuss what you perceive as strengths and limitations of your analysis._

## Conclusions
_What are the main take-home messages?_

_Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end_

>>>>

# Summary/Abstract

@@

The proposed study aims to assess the relationship between overall staffing rates with healthcare-associated infection rates.

First, we will provide a descriptive analysis of nursing homes across the United States including two different staffing measures, bed size, ownership, average number of residents per day. Second, we will utilize ANOVA methods to determine which, if any, average staffing characteristics are associated with the classifications of HAI rates as "better", "worse", or "as expected". Third, we will assess a variety of nursing home characteristics and selected staffing measures for associations with the risk adjusted rate of HAIs in the nursing home.

@@

## General Background Information

*Provide enough background on your topic that others can understand the why and how of your analysis*

## Description of data and data source

*Describe what the data is, what it contains, where it is from, etc. Eventually this might be part of a methods section.*

## Questions/Hypotheses to be addressed

*State the research questions you plan to answer with this analysis.*

@@

*Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement.*

## Data aquisition

*As applicable, explain where and how you got the data. If you directly import the data from an online source, you can combine this section with the next.*

## Data import and cleaning

*Write code that reads in the file and cleans it so it's ready for analysis. Since this will be fairly long code for most datasets, it might be a good idea to have it in one or several R scripts. If that is the case, explain here briefly what kind of cleaning/processing you do, and provide more details and well documented code somewhere (e.g. as supplement in a paper). All materials, including files that contain code, should be commented well so everyone can follow along.*

## Statistical analysis

*Explain anything related to your statistical analyses.*

@@

*Use a combination of text/tables/figures to explore and describe your data. Show the most important descriptive results here. Additional ones should go in the supplement. Even more can be in the R and Quarto files that are part of your project.*

@@

*To get some further insight into your data, if reasonable you could compute simple statistics (e.g. simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable. Though note that unless you pre-specified the outcome and main exposure, any "p\<0.05 means statistical significance" interpretation is not valid.*

@@

*Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc. This might again be code that is best placed in one or several separate R scripts that need to be well documented. You want the code to produce figures and data ready for display as tables, and save those. Then you load them here.*

@@

## Summary and Interpretation

*Summarize what you did, what you found and what it means.*

## Strengths and Limitations

*Discuss what you perceive as strengths and limitations of your analysis.*

## Conclusions

*What are the main take-home messages?*

*Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end*
